% Plotting FLR objects with ggplot2
% Iago Mosqueira, EC JRC - FLR Project
% 2013

```{r, setup, echo=FALSE, results='hide', message=FALSE}
library(knitr)
opts_chunk$set(dev='pdf', cache=TRUE)
options(width=50)
```

# Using ggplot2 with FLR objects
```{r, load, echo=FALSE, results='hide', message=FALSE}
library(ggplotFL)
data(ple4)
```

The `ggplot2` ^[\url{http://ggplot2.org/}] package provides a powerful alternative paradigm for creating simple and complex plots in R using the *Grammar of Graphics* ^[Wilkinson, L. 1999. *The Grammar of Graphics*, Springer. ISBN 0-387-98774-6.]

To facilitate the use of `ggplot2` methods in `FLR`, `ggplotFL` has been created. The main resources on offer in this package are overloaded versions of the `ggplot()` method that take directly certaing `FLR` classes, a new set of basic plots for some `FLR` classes, based on `ggplot2` instead of `lattice`, and some examples and documentation on how best make use of `ggplot2`'s powerful paradigm and implementation to obtain high quality plots for even farly complex data structures.

# The overloaded `ggplot` method

## FLQuant

```{r, flquant, echo=TRUE, tidy=TRUE, fig.cap="Combined line and point plot of a time series from an FLQuant object."}
ggplot(data=catch(ple4), aes(year, data)) + geom_point() + geom_line()
```

## FLQuants

## FLStock

# plot() for FLR classes

## FLStock
```{r, plotFLStock, echo=FALSE, cache=TRUE,  fig.cap="ggplot2 version of the standard plot() for FLStock, as applied to `ple4`"}
plot(ple4)
```
# Converting to data.frame

# Some extra examples

# More information

* The latest version of `ggplotFL` can always be installed using the `devtools` package, by calling
```{r, devtools, echo=TRUE, eval=FALSE}
	library(devtools)
	install_github('ggplotFL', 'flr')
```
* Learn about ggplot2
* reshape

# Version

* R: `r version$version.string`
* ggplot2: `r packageVersion('ggplot2')`
