% Plotting FLR objects with ggplot2 and ggplotFL
% Iago Mosqueira, EC JRC - FLR Project
% August 2013

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Plotting FLR objects with ggplot2 and ggplotFL}
-->

```{r, setup, echo=FALSE, results='hide', message=FALSE}
library(knitr)
opts_chunk$set(dev='pdf', cache=TRUE)
options(width=50)
```

# Using ggplot2 with FLR objects
```{r, load, echo=FALSE, results='hide', message=FALSE}
library(ggplotFL)
data(ple4)
set.seed(1423)
```

The `ggplot2` ^[\url{http://ggplot2.org/}] package provides a powerful alternative paradigm for creating simple and complex plots in R using the *Grammar of Graphics* ^[Wilkinson, L. 1999. *The Grammar of Graphics*, Springer. ISBN 0-387-98774-6.]

To facilitate the use of `ggplot2` methods in `FLR`, the `ggplotFL` package has been created. The main resources on offer in this package are overloaded versions of the `ggplot()` method that take directly certaing `FLR` classes, a new set of basic plots for some `FLR` classes, based on `ggplot2` instead of `lattice`, and some examples and documentation on how best make use of `ggplot2`'s powerful paradigm and implementation to obtain high quality plots for even fairly complex data structures.

# The overloaded `ggplot` method

## FLQuant

```{r, flquant, echo=TRUE, tidy=TRUE, fig.cap="Combined line and point plot of a time series from an FLQuant object."}
ggplot(data=catch(ple4), aes(year, data)) + geom_point() + geom_line()
```

## FLQuants

```{r, flquants, echo=TRUE, tidy=TRUE, fig.cap="Facet wrap line plot of some time series from an FLQuants object."}
ggplot(data=FLQuants(Yield=catch(ple4), SSB=ssb(ple4), F=fbar(ple4)), aes(year, data)) + geom_line() + facet_wrap(~qname, scales="free", nrow=3)
```
## FLStock

# New plot() method for FLR classes

The `ggplotFL` package also provides new versions of the `plot` method for a number of `FLR` classes. Each S4 class defined in any `FLR` package should have a `plot()` method defined that provides a visual summary of the contents of the object.

## FLStock
```{r, plotFLStock, echo=FALSE, cache=TRUE,  fig.cap="ggplot2 version of the standard plot() for FLStock, as applied to `ple4`"}
plot(ple4)
```

# Using ggplot2 by converting to data.frame

The methods shown above simply depend on conversion of `FLR` objects into `data.frame`, which can then be passed to `ggplot()`. Calling `ggplot` on an `FLR` object takes care of this conversion behind the scenes, but to obtain certains plots, it is best to directly convert the `FLR` objects into a `data.frame`.

# Examples

## EXAMPLE: plot quantiles of a simulation

To have full control over a plot of the median (or mean) and the confidence or probability intervals of a simulated or randomized time series, i.e. an `FLQuant` object with iters, we need to arrange the different values computed from the object in separate columns of a `data.frame`.

If we start with some random `FLQuant` object, such as
```{r, exsim1, echo=TRUE, cache=TRUE, fig.cap=""}
fla <- rlnorm(100, FLQuant(exp(cumsum(rnorm(25, 0, 0.1)))), 0.1) 
ggplot(fla, aes(factor(year), data)) + geom_boxplot() + xlab("")
```
we can first compute the necessary statistics on the object itself, as these operations are very efficient on an array. `quantile()` on an `FLQuant` will return the specified quantiles along the `iter` dimension. Let's extract the 10th, 25th, 50th, 75th and 90th quantiles.

```{r, exsim2, echo=TRUE, cache=TRUE}
flq <- quantile(fla, c(0.10, 0.25, 0.50, 0.75, 0.90))
```

The object can now be coerced to a `data.frame`
```{r, exsim3, echo=TRUE, cache=TRUE}
fdf <- as.data.frame(flq)
```
and inspected to see how the 100 `iters` have been now turned into the five requested quantiles

```{r, exsim4, echo=TRUE, cache=TRUE, results='markup'}
head(fdf)
```

The long format `data.frame` can be reshaped into a wide format one so that we can instruct `ggplot` to use the quantiles, now in separate columns, to provide limits for the shaded areas in `geom_ribbon`. To do this we can use `cast`, as follows

```{r, exsim5, echo=TRUE, cache=TRUE}
fdw <- cast(fdf, quant+year+unit+season+area~iter, value="data")
```

This creates a wide `data.frame` in which the `iter` column is spread into five columns named as the levels of its conversion into factor

```{r, exsim6, echo=TRUE, cache=TRUE}
levels(fdf[,'iter'])
```

We can now use those five quantile columns when plotting shaded areas using `geom_ribbon`. Please note that the column names returned by `quantile()` need to be quoted using backticks.

```{r, exsim7, echo=TRUE, cache=TRUE, fig.cap=""}
p <- ggplot(data=fdw, aes(x=year, y=`50%`)) + geom_line() +
	geom_ribbon(aes(x=year, ymin = `10%`, ymax = `90%`), fill="red", alpha = .15) +
	geom_ribbon(aes(x=year, ymin = `25%`, ymax = `75%`), fill="red", alpha = .25) +
	ylab("data")
print(p)
```

## EXAMPLE: Using FLQuants

Coercion using `as.data.frame`, combined with the use of `cast` and `melt` (from the `reshape` package), provides the `FLR` user with most tools required to create a large range of `ggplot`s out of any `FLR` object.

## EXAMPLE: Simulation trajectories plot


```{r, exspa, echo=TRUE, cache=TRUE, fig.cap=""}
fds  <- cast(as.data.frame(iter(fla, c(1, 4, 23, 56))), quant+year+unit+season+area~iter, value="data")

p + geom_line(data=fds, aes(year, c(`1`)), colour='red', size=1)

fds  <- as.data.frame(iter(fla, c(1, 4, 23, 56)))

p + geom_line(data=fds, aes(year, data, colour=iter), size=1) + theme(legend.position = "none")
```

## EXAMPLE: Bubble plots

```{r, exbub, echo=TRUE, cache=TRUE, fig.cap=""}
ggplot(catch.n(ple4), aes(year, as.factor(age), size=data)) + geom_point(shape=21) + scale_size(range = c(1, 20)) + ylab("age") + theme(legend.position = "none")
```

## EXAMPLE: Residual plots
```{r, exres, echo=TRUE, cache=TRUE, fig.cap=""}
dat <- as.data.frame(catch.n(ple4))
dat$resid <- dat$data - mean(dat$data)

ggplot(dat, aes(year, as.factor(age), size=resid)) +
	geom_point(shape=21, aes(colour=factor(sign(resid)), fill=factor(sign(resid)))) +
	scale_size(range = c(1, 20)) +
	scale_colour_manual(values=c("black", "white")) +
	scale_fill_manual(values=c("lightgray", "black")) +
	theme(legend.position = "none") + ylab("age")
```

## EXAMPLE: Cohort

```{r, excoh, echo=TRUE, cache=TRUE, fig.cap=""}
ggplot(catch.n(ple4), aes(cohort, data)) + geom_line(aes(group=age, colour=factor(age))) + scale_colour_hue()
```

# More information

* The latest version of `ggplotFL` can always be installed using the `devtools` package, by calling
```{r, devtools, echo=TRUE, eval=FALSE}
	library(devtools)
	install_github('ggplotFL', 'flr')
```
* Any bug, question or suggestion on `ggplotFL` can be notified to the developers at the `ggplotFL` issue page ^[https://github.com/flr/ggplotFL/issues], or on the `FLR` mailing list.

* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage \marginnote{http://flr-project.org}.

* To learn more about ggplot2, visit the ggplot2 website ^[\url{http://ggplot2.org/}], or read the ggplot2 book.^[Wickham, H. 2009. *ggplot2: Elegant Graphics for Data Analysis*. Springer, Use R! Series. ISBN 978-0-387-98140-6]

## Package versions

* R: `r version$version.string`
* ggplot2: `r packageVersion('ggplot2')`
* ggplotFL: `r packageVersion('ggplotFL')`
* FLCore: `r packageVersion('FLCore')`
